---
title: "eda"
format: html
editor: visual
---

# How many cases

pending cases are codded as tapeyear == 2099 for some reason.

```{r}

library(DBI)
library(tidyverse)
source("lookups.R")

ds <- arrow::open_dataset("data")

ds |> 
  group_by(TAPEYEAR) |> 
  tally() |> 
  collect() -> temp

temp |> 
  filter(TAPEYEAR != 2099) |> 
  arrange(TAPEYEAR) |> plot()

```

Closure month by disposition

```{r}

monthly_disposition <- ds |> 
  arrow::to_duckdb() |> 
  mutate(closed_month = lubridate::floor_date(TERMDATE, "month")) |> 
  group_by(closed_month, DISP) |> 
  tally() |> 
  collect()

monthly_disposition |> 
  filter(DISP != -8) |> 
  filter(closed_month >= lubridate::ymd("1988-01-01")) |> 
  left_join(disp) |> 
  ggplot(aes(closed_month, n, group = type, color = as.factor(type))) +
  geom_line() 
  
```

Basis of Jurisdiction in federal courts

Federal Plaintiff litigation

```{r}

monthly_juris <- ds |> 
  arrow::to_duckdb() |> 
  mutate(closed_month = lubridate::floor_date(TERMDATE, "month")) |> 
  group_by(closed_month, JURIS) |> 
  tally() |> 
  collect()

monthly_juris |> 
  filter(JURIS == 1) |> 
  filter(closed_month >= lubridate::ymd("1988-01-01")) |> 
  ggplot(aes(closed_month, n)) +
  geom_line() + 
  labs(
    x = "Date closed"
    , y = "Total Federal Suits"
    , title = "Federal lawsuits with US government Plaintiff")

```

Looking at the "nature of the suit"

```{r}

ds |> 
  filter(JURIS == 1) |> 
  filter(TAPEYEAR != 2099) |> 
  group_by(TAPEYEAR, NOS) |> 
  tally() |> 
  collect() |>  
  ggplot(aes(TAPEYEAR, n, color = as.factor(NOS))) +
  geom_line()

```

```{r}

ds |> 
  filter(JURIS == 1) |> 
  filter(NOS == 152) |> 
  filter(TAPEYEAR != 2099) |>
  arrow::to_duckdb() |> 
  mutate(closed_month = lubridate::floor_date(TERMDATE, "month")) |> 
  group_by(closed_month) |> 
  tally() |> 
  collect() |>  
  ggplot(aes(closed_month, n )) +
  geom_line() +
  labs(x = "closed month", y = "Number of cases", title = "Student loan recovery lawsuits filed by the US government")  


```

how long was a case open when it closed

```{r}

ds |> 
  head(1000) |> 
  pull(FILEDATE) 

ds |> 
  arrow::to_duckdb() |> 
  filter(DISP != -8) |> 
  mutate(
    days_to_resolve =  TERMDATE - FILEDATE
    , month = lubridate::floor_date(TERMDATE, "year")) |>
  group_by(month) |> 
  summarise(days = mean(days_to_resolve)) |> 
  collect() -> days_ts
      
days_ts |> 
  filter(month >= lubridate::ymd("1988-01-01")) |> 
  ggplot(aes(month, days)) +
  geom_line() +
  labs(title = "mean days to resolve or transfer")

ds |> 
  filter(DISP == -8
         , JURIS == 4
         , FILEDATE > "2000-01-01") |> 
  arrow::to_duckdb() |> 
  mutate(month = lubridate::floor_date(FILEDATE, "month")) |> 
  group_by(month, JURIS) |> 
  tally() |> 
  ggplot(aes(month, n, color = as.factor(JURIS))) + 
  geom_line() 
  

```

so I think we want to model the time to closure with a censored model.

let's pick a type of jurisdiction (federal question) and a couple of circuits (5th and 11th) over the last 7 years.

```{r}

table_df <- function(df, column){
  df |> 
    group_by({{column}}) |> 
    tally() |> 
    collect()
}

ds |> 
  mutate(full_court = paste0(CIRCUIT, "-", DISTRICT )) |> 
  group_by(full_court,  TERMDATE ) |> 
  tally() |> 
  collect()-> ts


ts |> 
  filter(TERMDATE >= "2020-01-01") |> 
  mutate(weekly = lubridate::floor_date(TERMDATE, unit = "week")) |> 
  filter(stringr::str_starts(full_court, "1-")) |> 
  group_by(full_court, weekly) |> 
  summarise(n = sum(n)) |> 
  ggplot(aes(weekly, n, color = full_court)) +
  geom_line() 

```

```{r}

years <- ds |>
  group_by(TAPEYEAR) |> 
  tally() |> 
  collect()

res <- list()

for(year in seq_along(years$TAPEYEAR)) {
  res[[year]] <- ds |> 
    arrow::to_duckdb() |> 
    filter(TAPEYEAR == !!years$TAPEYEAR[[year]] ) |> 
    filter(CIRCUIT == 5) |> 
    collect() 
  }

df <- reduce(res, bind_rows)

df <- df |> 
  mutate(
    censored = ifelse(TERMDATE == "1900-01-01"
                      , "censored"
                      , "not censored")
    , td_updated = ifelse(TERMDATE == "1900-01-01"
                        , lubridate::today()
                        , TERMDATE) |> as_date()
    , top_coded = ifelse(DEMANDED == 9999, TRUE, FALSE) 
    , demand_scaled = scale(DEMANDED * 1000)[,1]
    , time_to_resolve = td_updated - FILEDATE + 1
    , ttr_log = log(as.integer(time_to_resolve))
    , day_of_filing = FILEDATE - min(FILEDATE)
    )


df <- df |> 
  group_by(DISTRICT, FILEDATE)  |> 
  mutate(n = n()) |> 
  group_by(DISTRICT, td_updated) |> 
  mutate(n_closed = n()) |> 
  group_by(DISTRICT, TRANSDAT) |>
  mutate(n_transf_in = n()) |> 
  mutate(n_transf_in = ifelse(is.na(TRANSDAT), 0, TRANSDAT)) |> 
  mutate(across(starts_with("n"), \(x) ifelse(is.na(x), 0, x) )) |> 
  group_by(DISTRICT) |> 
  arrange(FILEDATE) |> 
  mutate(
    net = (n + n_transf_in) - n_closed
    , cum_n = cumsum(as.numeric(net))
  ) 

df |> 
  filter(DISTRICT == 40) |> 
  ggplot(aes( FILEDATE, cum_n, color = DISTRICT)) +
  geom_line()

temp <- df |> filter(DISTRICT == 40)

temp |> select(FILEDATE, TERMDATE, censored, TRANSDAT, n, n_transf_in, n_closed, net, cum_n) |> 
  arrange(FILEDATE) -> temp 

view(temp)

```

I think we maybe want to keep track of the number of open cases in a district court?

And then we can use that, plus the number of cases filed in same week as your case and plus demand amount?

```{r}

x <- c(0, 4, 8)
y <- c(0, 8, 0)

beas_tbl <- bind_cols(
  x = x
  ,y = y )


beas_tbl |> 
  ggplot(aes(x, y)) + 
  geom_area(color = "red", size = 9, fill = "red")
```
